---
title: "Lab2_assignment"
author: "Amelia Marcantonio-Fields"
date: "9/13/2021"
output: github_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup_package, warning = FALSE, message = FALSE}
library(viridis)
library(tidyverse)
library(tidycensus)
library(dplyr)
library(sf)
library(tmap) # mapping, install if you don't have it
set.seed(717)
```

This assignment if for you to complete a short version of the lab notes, but you have to complete a number of the steps yourself.
You will then knit this to a markdown (not an HTML) and push it to your GitHub repo.
Unlike HTML, the RMarkdown knit to `github_document` can be viewed directly on GitHub.
You will them email your lab instructor with a link to your repo.

Steps in this assignment:

1.  Make sure you have successfully read, run, and learned from the `MUSA_508_Lab2_sf.Rmd` Rmarkdown

2.  Find two new variables from the 2019 ACS data to load.
    Use `vars <- load_variables(2019, "acs5")` and `View(vars)` to see all of the variable from that ACS.
    Note that you should not pick something really obscure like count_38yo_cabinetmakers because you will get lots of NAs.
    
```{r acs_vars, cache = FALSE, message = FALSE, warning = FALSE, results=FALSE}

census_api_key("9a4eb0449cd787a44a9402bccce6e20769078384",overwrite = TRUE)

vars <- load_variables(2019, "acs5")

view(vars)

lab2_acs_vars <- c("B01001_001E", # ACS total Pop Estimate
                    "B16010_041E", # Total Bachelor's Degree or Higher
                    "B19019_001E") # Median HH Income (2019 $)

myTracts2 <- c("42101000803",
               "42101000804",
               "42101001202")

acsTractsPHL.2019.sf <- get_acs(geography = "tract",
                             year = 2019, 
                             variables = lab2_acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(lab2_acs_vars)) %>%
  rename (total_pop.2019 = B01001_001E,
          total_BDorhigher.2019 = B16010_041E,
          med_HH_income.2019 = B19019_001E) %>%
  mutate(pctBDorhigher.2019 = total_BDorhigher.2019/total_pop.2019) %>%
  mutate(Rittenhouse = ifelse(GEOID %in% myTracts2, "RITTENHOUSE", "REST OF PHILADELPHIA"))
```


3.  Pick a neighborhood of the City to map.
    You will need to do some googling to figure this out.
    Use the [PHL Track Explorer](https://data-phl.opendata.arcgis.com/datasets/census-tracts-2010/explore?location=40.002759%2C-75.119097%2C11.91) to get the `GEOID10` number from each parcel and add them to the `myTracts` object below.
    This is just like what was done in the exercise, but with a different neighborhood of your choice.
    Remember that all GEOIDs need to be 10-characters long.


4.  In the first code chunk you will do that above and then edit the call-outs in the dplyr pipe sequence to `rename` and `mutate` your data.

5.  You will transform the data to `WGS84` by adding the correct EPSG code.
    This is discussed heavily in the exercise.

```{r}
acsTractsPHL.2019.sf_WGS84 <- acsTractsPHL.2019.sf %>% 
  st_transform(crs = "EPSG:4326")

st_crs(acsTractsPHL.2019.sf_WGS84)
```

6.  You will produce a map of one of the variables you picked and highlight the neighborhood you picked.
    There are call-out within the `ggplot` code for you to edit.
    
```{r ggplot_geom_sf, warning = FALSE, echo = FALSE}

ggplot()+
  geom_sf(data = acsTractsPHL.2019.sf, aes(fill = pctBDorhigher.2019),
          color = "transparent")+
  geom_sf(data = acsTractsPHL.2019.sf_WGS84 %>%
            filter(Rittenhouse == "RITTENHOUSE") %>%
            st_union(),
          color = "white",
          fill = "transparent")+
 scale_fill_viridis(
          alpha = 1,
          begin = 1,
          end = 0,
          direction = 1,
          discrete = FALSE,
          option = "F",
          labels = scales::percent,
          name = "% of Population", 
          limits = c(0, 1)) +
 # Bonus to figure out different color ramps with scale_fill_viridis()
  labs(
    x = "Longitude", y = "Latitude", fill = "",
    title = "Bachelor's Degree or Higher by Census Tract",
    subtitle = "",
    caption = "Data: US Census Bureau, ACS 2019 5-year estimates")
```

7.  You can run the code chunks and lines of code as you edit to make sure everything works.

8.  Once you are done, hit the `knit` button at the top of the script window (little blue knitting ball) and you will see the output.
    Once it is what you want...

9.  Use the `Git` tab on the bottom left of right (depending on hour your Rstudio is laid out) and click the check box to `stage` all of your changes, write a commit note, hit the `commit` button, and then the `Push` button to push it to Github.

10. Check your Github repo to see you work in the cloud.

11. Email your lab instructor with a link!

12. Congrats!
    You made a map in code!